---
title: "Comparison of multiple distributions"
author: "Fahad Deshmukh"
date: "01 08 2023"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)

```

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
# Read data
df <- read.csv("D:/multiple-distributions/Data/SwimmingTimes.csv")
```

```{r}
str(df)
```

```{r}
# Change categorical variables to factors
df$Name <- as.factor(df$Name)
df$Category <- as.factor(df$Category)
str(df)
```

```{r}
# Summary statistics
summary(df)
```

```{r}
# Get the maximum number of Name count
max(table(df$Name))
```

```{r}
# Swimmers which competed in two different categories
tbl <- as.data.frame(table(df$Name))
bi_swimmmers <- tbl[tbl$Freq>1,]$Var1

# Subset dataframe to get only the data of swimmers which competed twice
bi_swimmmers_df <- df[df$Name %in% bi_swimmmers,]
bi_swimmmers_df[order(bi_swimmmers_df$Name),]
```

```{r}
# As there are some swimmers who competed in two different categories we will remove them from one competition. By doing so we will have each swimmer only once in the data set.   
df <- df[-c(2,5,34,19,42,51,79),]
```

```{r}
# New Category sizes
summary(df)
```


```{r}
# Histogram
hist(df$Time, breaks = 50, main = "", xlab = "Time (seconds)", ylab="Frequency", cex.lab = 1.5)
# Choice of a good width?
# Label Size?
```

```{r}
# Histogram for each Category
library(plyr)
mu <- ddply(df, "Category", summarise, grp.mean=mean(Time))

p <- ggplot(df, aes(x=Time)) +
  geom_histogram(color="black", fill="white", binwidth = 0.4) +
  facet_grid(Category ~ ., scales="free_y") +
  theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=16),
        legend.text=element_text(size=12)) +
  theme(legend.position='none') +
  labs(x="Time (seconds)", y = "Frequency") 
  # Add mean lines
  p + geom_vline(data=mu, aes(xintercept=grp.mean, color="red"), linetype="solid")
```


```{r}
# Boxplot of each category
df %>% 
  ggplot(aes(factor(Category), Time, fill = factor(Category))) +
  geom_boxplot() +
  theme(panel.spacing = unit(0, "lines"),
        panel.border = element_rect(fill = NA),
        strip.background = element_blank(),
        #axis.title.x = element_blank(),
        legend.position = "none",
        strip.placement = "outside") +
  theme(panel.background = element_blank()) +
  theme(axis.line = element_line(colour = "black")) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
  theme(legend.position='none') +
  labs(y="Time (seconds)", x = "Category") +
  theme(axis.text=element_text(size=12), 
        axis.title=element_text(size=16),
        legend.text=element_text(size=12))

# Color? Should we use black (the same color) for all boxes?
```

```{r}
# Mean of each category
hm <- setNames(aggregate(df$Time, by=list(df$Category), FUN=mean), c("Category", "Mean")) 

# Variance of each category
hm$Variance <- tapply(df$Time, df$Category, FUN = var)

# Standard Deviation of each category
hm$`Standard Deviation` <- tapply(df$Time, df$Category, FUN = sd)

hm 
```


## Assumption (Normality)

```{r}
# qqplots for normality check
ggplot(df, aes(sample = Time)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~Category, scales = "free") +
  labs(y="Sample", x = "Theoretical")
```

Since the sample values generally match the theoretical values, we conclude that the time values in each category are normally distributed. Therefore, we can say that the values in each group are normally distributed and the normality assumption is satisfied. 

```{r}
# !!! OPTIONAL !!!
#generate vector of 15 values that follow standard normal distribution
data <- rnorm(15)
data <- as.data.frame(data)
ggplot(data, aes(sample = data)) +
  stat_qq() +
  stat_qq_line()
```

We see that simulated normally distributed random values have similar pattern as the ones obtained with the sample values.

```{r}
# !!! not included in analysis (optional) !!!
# Shapiro test for normality check
shapiro.test(df[df$Category=="Backstroke",]$Time)
shapiro.test(df[df$Category=="Breaststroke",]$Time)
shapiro.test(df[df$Category=="Butterfly",]$Time)
shapiro.test(df[df$Category=="Freestyle",]$Time)
shapiro.test(df[df$Category=="Medley",]$Time)
```

# Assumption (Homogeneity)

```{r}
# Standard Deviation of each category
hm <- setNames(aggregate(df$Time, by=list(df$Category), FUN=sd), c("Category", "Standard Deviation")) 

# Variance of each category
hm$Variance <- tapply(df$Time, df$Category, FUN = var)

hm
```

The variance and the standard deviation of each category is similar. The category Butterfly have higher variability but the difference is not big. Hence, we can say that the variance of the categories are the same and the homogeneity assumption is satisfied.

```{r, message=FALSE}
# !!! OPTIONAL !!!
library(car)
# Levene's test with one independent variable
leveneTest(Time ~ Category, data = df)
```


# Assumption (Independence)

It was actually violated if had used the whole data set, as there were the same swimmers in different groups. To handle this issue we removed the 7 observations where the same swimmer competed in another category as well. The selection of the observations are made in a way to keep the group sizes close to each other. There were 6 swimmers who competed in medley in addition to another category. This was taken into consideration. 

As all three assumptions are satisfied (Normality, Homogeneity, Independence) we can use the One-way ANOVA test.

# Parametric Test (One-way ANOVA)

```{r}
# One-way ANNOVA
res_aov <- aov(Time ~ Category, data = df)
summary(res_aov)
```

The obtained p-value is significantly small at the significance level $alpha$ = 0.05. Hence, we reject the null hypothesis (All categories have the same mean value), and we conclude that at least one category has significantly different mean value than others. Now, we will check the pairwise differences using pairwise t-tests.

First, we do pairwise t-tests without any adjustment for multiple testing.

```{r}
# No adjustment
pairwise.t.test(df$Time, df$Category, p.adjust.method = "none")
```

Except the pair Butterfly-Backstroke (0.516), all pairwise tests result is significantly (at 0.05 significance level) small p-values. Except Butterfly-Backstroke all category pairs have significantly different mean values. However for the Butterfly-Backstroke we fail to reject the null hypothesis (Butterfly and Backstroke categories have the same mean value) and conclude that there is no significant evidence to say these two categories have different mean values. 

However, as we use the same data set for multiple tests we need to adjust the testing procedure accordingly. For this purpose we use two different adjustment methods, "bonferroni" and "holm". 

```{r}
# "bonferroni" adjustment (more conservative)
pairwise.t.test(df$Time, df$Category, p.adjust.method = "bonferroni")
```

With bonferroni adjustment method we obtain p-values which are still significantly small (alpha = 0.05) for all pairs except Butterfly-Backstroke (1.0000). Hence, the conclusions remains the same. Now, we use the "holm" method for adjustment which is less conservative than the "bonferroni" method.

```{r}
# "holm" adjustment ->  less conservative
pairwise.t.test(df$Time, df$Category, p.adjust.method = "holm")
```

We again obtain significantly (alpha = 0.05) small p-values for all pairs except Butterfly-Backstroke (0.5165). The conclusions remain the same. We fail to reject the null hypothesis (The mean values are the same) for the pair Butterfly-Backstroke and say there is no enough evidence to say both categories are different. However, for all other pairs we reject the null hypothesis and conclude that the categories in the pairs are significantly different from each other.
```{r}
df.Backstroke<-filter(df,Category=='Backstroke')
df.Breaststroke<-filter(df,Category=='Breaststroke')
df.Butterfly<-filter(df,Category=='Butterfly')
df.Freestyle<-filter(df,Category=='Freestyle')
df.Medley<-filter(df,Category=='Medley')
```
```{r}
summary(df.Backstroke)
IQR(df.Backstroke$Time)
```
```{r}
summary(df.Breaststroke)

```

```{r}
summary(df.Butterfly)
IQR(df.Butterfly$Time)
```

```{r}
summary(df.Freestyle)
```

```{r}
summary(df.Medley)
```

